---
title: "LightCurveAnalysis"
author: "Doug Ratay"
date: "August 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("imageFunctions.R")
```

# Thinking about Light Curves

We're going to generate some light curves with a Gaussian Process, create some DMDT images, and then do some classification.  This is an intellectual excersize to see where it goes.

## Generate Truth Signals

We're going to pick some basic functions to serve as our "truth" data.  There's no reason behind the selection of these functions other than that they're sort of similarly shaped and scaled and with enough noise there might be some overlap.

```{r}
require(tibble)
require(dplyr)
require(magrittr)
require(ggplot2)
require(tidyr)
require(purrr)

typeNames = c('truth1','truth2','truth3','truth4')
timeVals = 0:100
numTimes = length(timeVals)
functionList = list(function(x){sin(x/35)},
                    function(x){1.5*sin(x/35)},
                    function(x){sin(x/35)+0.01*x},
                    function(x){sqrt(x)/10 + 0.01*x})

truthTable <- map2(.x=typeNames,.y=functionList,.f=function(name,func) {
  
  initTable <- tibble(time=timeVals)
  initTable %>% mutate(!!name := func(time)) %>% gather(key='func',value='value',-time)
}) %>% bind_rows()

truthTable %>% ggplot(aes(x=time,y=value,color=func)) + geom_point() + geom_line()
```

To make things a bit more intersting with noise and overlap, we purposesly lose track of the original functions, subsample the points a bit, and fit Gaussian Processes.  This will yield different noise values at different times and make things a bit more difficult.

```{r}

numToKeep = 50
keptPoints <- map(.x=typeNames,.f=function(name) {
  tibble(func = rep(name,numToKeep),point = sort(sample(timeVals,numToKeep)))
}) %>% bind_rows()


sampledTable <- map(.x=typeNames,.f=function(name) {
  
  indices = keptPoints %>% filter(func==name) %>% select(point)
  originalSamples = truthTable %>% filter(func==name)
  originalSamples %>% filter(time %in% as.array(indices$point))
}) %>% bind_rows()

sampledTable %>% ggplot(aes(x=time,y=value,color=func)) + geom_point() + geom_line()

#Add some noice to make things interesting

sampledTable %<>% mutate(noiseValue = value + rnorm(n(),mean=0,sd=0.01))

sampledTable %>% ggplot(aes(x=time,y=noiseValue,color=func)) + geom_point()

```

Now we fit with a Process

```{r}

listOfGPs <- typeNames %>% map(.f=function(name) {
  
  sampledTable %>% filter(func==name) %>% transmute(time=time,value=noiseValue) %>% learnGP()
})
```

```{r}
#make some plots of the structure of the GP

predictions <- map2(.x=listOfGPs,.y=typeNames,.f=function(gp,name) {
  
  prediction <- gp$predict(timeVals,se=TRUE)
  prediction %<>% mutate(time=timeVals,upper=mean+2*se,lower=mean-2*se,type=rep(name,numTimes))
}) %>% bind_rows()
```

```{r}
predictions %>% ggplot(aes(x=time,y=mean,color=type)) + geom_line() + geom_linerange(aes(ymin=lower,ymax=upper)) 

```

## Generate Samples for Learning

With the GPs, we can create sample curves.  We'll start with the easiest case where we have observations over the full time range.  We'll have a range in the number of observations, that will be randomized, with a minimum of 10 observations and a maximum of 30.

```{r}

numTruthProfilesPerType = 30
numPointsPerProfileRange = 10:30
numObsPerSample = sample(numPointsPerProfileRange,numTruthProfilesPerType,replace=TRUE)

observationTimeList = 1:numTruthProfilesPerType %>% map(.f=function(index) {
  sort(sample(timeVals,size = numObsPerSample[index]))
})

observationTimes <- 
  tibble(index=1:numTruthProfilesPerType,samplePoints=observationTimeList)

observationList <- map2(.x=listOfGPs,.y=typeNames,.f=function(gp,name) {
  
  samples <- generateGPSamples(gp,observationTimes)
  samples %>% mutate(type=rep(name,n()))
}) %>% bind_rows()

```

```{r}

#Lets do this with gganimate when connected to the internet again.

1:length(typeNames) %>% walk(.f=function(index) {
  
  limitedData <- observationList %>% filter(type==typeNames[index]) %>% mutate(trial=as.factor(trial))
  p <- limitedData %>% ggplot(aes(x=time,y=value,color=trial)) + geom_point()
  print(p)
})

```

## Delvelop DLDT Images

Now we convert these samples into our DLDT images.  First we take the differences.

```{r}

#Expect the input table to the function to have columns = time,value

# separate by type
dldtData <- 1:length(typeNames) %>% map(.f= function(typeIndex) {
  
  singleType <- observationList %>% filter(type == typeNames[typeIndex])
  #and then separate by trial
  dldts <- 1:numTruthProfilesPerType %>% map(.f=function(trialIndex){
    
    inputTable <- singleType %>% filter(trial==trialIndex) %>% select(time,value)
    calculateDValDTime(inputTable) %>% mutate(trial = rep(trialIndex,n()))    
  }) %>% bind_rows()
  
  dldts %>% mutate(type = rep(typeNames[typeIndex],n()))
}) %>% bind_rows()

```

```{r}

#look at this later with some annimation or faceting

# dldtData %>% filter(type=='truth1') %>% ggplot(aes(x=time,y=value,color=trial)) + geom_point()

1:length(typeNames) %>% walk(.f=function(typeIndex){
  
  reducedData <- dldtData %>% filter(type==typeNames[typeIndex])
  p <- reducedData %>% ggplot(aes(x=time,y=value,color=trial)) + geom_point() + ylim(c(-2,2))
  print(p)
})

```


We could have done both the differences and imaging all at once, but we did them separately to visualize and we're not really dealing with that much data here.  Now to the images.

```{r}
#We have to set up our bin values in the way required by ash::bin2
#The limts value is a matrix with x on the first row and y on the second
limits <- matrix(c(0,-2,100,2),2,2)
#The bins value is a 2 element array with the number of xbins first, ybins second
#We'll make a 16x16 image just because
xPixels = 16
yPixels = 16
bins <- c(xPixels,yPixels)
numPixels = xPixels * yPixels

#We do the same loops as above
imageListOverType <- 1:length(typeNames) %>% map(.f= function(typeIndex) {
  
  singleType <- dldtData %>% filter(type == typeNames[typeIndex])
  #and then separate by trial
  imageListOverTrial <- 1:numTruthProfilesPerType %>% map(.f=function(trialIndex){
    
    inputTable <- singleType %>% filter(trial==trialIndex) %>% select(time,value)
    imageFrame <- binDValDTime(inputTable,limits,bins) %>% as.data.frame() 
    imageFrame %>% mutate(trial=rep(trialIndex,n()))     
  }) %>% bind_rows() 
  
  imageListOverTrial %>% mutate(type = rep(typeNames[typeIndex],n()))
}) %>% bind_rows()

```

```{r}

1:length(typeNames) %>% walk(.f=function(typeIndex) {
  
  p <- imageListOverType %>% filter(type==typeNames[typeIndex]) %>%  ggplot(aes(x,y))+geom_raster(aes(fill=value)) + facet_wrap(vars(trial))
  print(p)
  
})

# imageListOverType %>% filter(type=='truth4',trial==1) %>% select(x,y,value) %>% ggplot(aes(x,y))+geom_raster(aes(fill=value))

```

## Analyze Images

The nice thing about converting to the images is that we now have a regular size across all samples to either do visualization or comparisons.  Here we'll take a slight detour and put all the points into tsne and see what happens.

```{r}
require(ggforce)

#most of what we have to do here is organize the input data.  Although that will happen in the function.

allImagesInRows <- widenImageDataFrame(imageListOverType,numPixels)
tsneOutput <- getTSNEEmbedding(allImagesInRows,numPixels)
```

```{r}
tsneOutput %>% ggplot(aes(x=V1,y=V2,color=type)) + geom_point()

#try a second one with some ellipses
tsneOutput %>% ggplot(aes(x=V1,y=V2,color=type)) + geom_mark_ellipse(aes(fill=type)) + geom_point()

ggplot(tsneOutput, aes(x=V1,y=V2)) +
  geom_delaunay_tile(alpha = 0.3) + 
  geom_delaunay_segment2(aes(colour = type, group = -1), size = 1,
                         lineend = 'round')



```

## Learn Classifier

Having made some pictures, we begin to learn a model.  We've already widened the data into a matrix that is appropriate for using in an NN.  We attempt that here.

```{r}

#A function accomplishes all of the work of training the model, and knows something about the structure we want to use, so we send in our data
require(keras)
require(tensorflow)
#install_keras()

kerasModel <- trainKerasModel(allImagesInRows)

```

```{r}
#We'll test our training data here.
predictionTable <- testKerasModel(allImagesInRows,kerasModel)

```

```{r}
predictionTable %>% ggplot(aes(correct)) + geom_bar()

```




